---
title: "Assignment 1"
author: "Danielle Wu"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Assignment 1

```{r libraries}
library (lsa)
library(tm)
library(rtweet)
library(wordcloud)
library(ggplot2)
library(tokenizers)
library(plyr)
library(Rgraphviz)
```
* load libraries

```{r data, echo=FALSE}
tweets_set1 <- read.csv("TADA_A1_set1.csv")
tweets_set2 <- read.csv("TADA_A1_set2.csv")

tweet_text1 <- tweets_set1$text
tweet_text2 <- tweets_set2$text

tweetall <- rbind(tweets_set1,tweets_set2)
```
* load data

```{r summary, echo=FALSE}
#file information, total posts
summary(tweets_set1)
summary(tweets_set2)
```
* view csv contents

```{r date, echo=FALSE}
#date range for tweets
min(tweets_set1$date)
max(tweets_set1$date)

min(tweets_set2$date)
max(tweets_set2$date)
tweets_set2 <- tweets_set2[-c(177826),]
max(tweets_set2$date)
```
*remove bad row and get date range 

```{r lang, echo=FALSE}
unique(tweetall$lang)

count(tweetall,'lang')
```
*get languages

```{r grep, echo=FALSE}
pat <- c("methadone|suboxone|fentanyl|fentanil|methodone|suboxen|suboxin|methadose|dolophine|actiq|duragesic|fentora")
sum(grepl(pat, tweets_set1$text,ignore.case=TRUE))
sum(grepl(pat, tweets_set2$text,ignore.case=TRUE))
```
* tweets mentioning methadone, suboxone, or fentanyl

```{r suboxone, echo=FALSE}
sub <-c("suboxone|suboxen|suboxin|buprenorphine-naloxone")

sum(grepl(sub, tweets_set1$text,ignore.case=TRUE))
sum(grepl(sub, tweets_set2$text,ignore.case=TRUE))
```
* tweets mentioning suboxone

```{r methadone, echo=FALSE}
methadone <- ("methadone|methodone|methadose|dolophine")

sum(grepl(methadone, tweets_set1$text,ignore.case=TRUE))
sum(grepl(methadone, tweets_set2$text,ignore.case=TRUE))
```
* tweets mentioning methadon

```{r fent, echo=FALSE}
fentanyl <- ("fentanyl|fentanil|actiq|duragesic|fentora")
sum(grepl(fentanyl, tweets_set1$text,ignore.case=TRUE))
sum(grepl(fentanyl, tweets_set2$text,ignore.case=TRUE))
```
* tweets mentioning fentanyl 

```{r analogs, echo=FALSE}
fent_text1 <- tweet_text1[c(grep(fentanyl, tweet_text1,ignore.case=TRUE))]
fent_text2 <- tweet_text2[c(grep(fentanyl, tweet_text2,ignore.case=TRUE))]

fent_text <- append(fent_text1,fent_text2)


analogs <- c("acetylfent|butyrfent|carfent|alfent|sufent|remifent|isofent|furanylfent|methylfent|acryloylfent|alfent|flurofent|furanylfent|fluorofent")
sum(grepl(analogs, fent_text,ignore.case=TRUE))
```
* tweets mentioning fentanyl analogs

# Subxone Tweet Set 
```{r subset1, echo=FALSE}
sub_text1 <- tweet_text1[c(grep(sub, tweet_text1,ignore.case=TRUE))]
sub_text2 <- tweet_text2[c(grep(sub, tweet_text2,ignore.case=TRUE))]

sub_text <- append(sub_text1,sub_text2)

```
* load suboxone related tweets

```{r preprocesssub1, echo=FALSE}
subcorpus_corpus1 <- VCorpus(VectorSource(sub_text))

#preprocessing: lowercase
subcorpus_corpus1 <- tm_map(subcorpus_corpus1, content_transformer(tolower))
#remove stop words 
subcorpus_corpus1 <- tm_map(subcorpus_corpus1, removeWords, stopwords("en"))
subcorpus_corpus1 <- tm_map(subcorpus_corpus1, removeWords, stopwords("ca"))
subcorpus_corpus1 <- tm_map(subcorpus_corpus1, removeWords, stopwords("es"))
subcorpus_corpus1 <- tm_map(subcorpus_corpus1, removeWords, stopwords("fr"))
#preprocessing: remove punctuation
subcorpus_corpus1 <- tm_map(subcorpus_corpus1, removePunctuation)
#preprocessing:remove numbers
subcorpus_corpus1 <- tm_map(subcorpus_corpus1, removeNumbers)
#preprocessing: remove stems
subcorpus_corpus1 <- tm_map(subcorpus_corpus1, stemDocument)
```
* suboxone-set 1 preprocess

```{r dtmsub1, echo=FALSE}
dtmsub1<-removeSparseTerms(DocumentTermMatrix(subcorpus_corpus1),0.995)
inspect(dtmsub1)
freqterms <- findFreqTerms(dtmsub1, lowfreq = 950)
print(freqterms)

findAssocs(dtmsub1,"suboxon",0.04)
```
* create document term matrix
* find associations

```{r sub1_new, echo=FALSE}
myStopwords <- c("suboxon","suboxin","suboxen","buprenorphinenaloxon")
subcorpus_corpus1_new <- tm_map(subcorpus_corpus1, removeWords, myStopwords)

wordcloud(subcorpus_corpus1_new, min.freq=800, max.words=15, scale=c(3,.1), random.order=FALSE, colors=brewer.pal(12, "Set3"))
```
* add new suboxone stop words
* generate word cloud


# Methadone Tweet Set 1
```{r methtext1, echo=FALSE}
meth_text1 <- tweet_text1[c(grep(methadone, tweet_text1,ignore.case=TRUE))]
meth_text2 <- tweet_text2[c(grep(methadone, tweet_text2,ignore.case=TRUE))]

meth_text <- append(meth_text1,meth_text2)
```
* load methadone tweets

```{r premeth, echo=FALSE}
#preprocess subtext
meth_corpus1 <- VCorpus(VectorSource(meth_text))

#preprocessing: lowercase
meth_corpus1 <- tm_map(meth_corpus1, content_transformer(tolower))
#remove stop words 
meth_corpus1 <- tm_map(meth_corpus1, removeWords, stopwords("en"))
meth_corpus1 <- tm_map(meth_corpus1, removeWords, stopwords("ca"))
meth_corpus1 <- tm_map(meth_corpus1, removeWords, stopwords("es"))
meth_corpus1 <- tm_map(meth_corpus1, removeWords, stopwords("fr"))

#preprocessing: remove punctuation
meth_corpus1 <- tm_map(meth_corpus1, removePunctuation)
#preprocessing:remove numbers
meth_corpus1 <- tm_map(meth_corpus1, removeNumbers)
#preprocessing: remove stems
meth_corpus1 <- tm_map(meth_corpus1, stemDocument)
```
* preprocessing

```{r dtmmeth1, echo=FALSE}
dtmmeth1 <-removeSparseTerms(DocumentTermMatrix(meth_corpus1),0.995)
inspect(dtmmeth1)
findAssocs(dtmmeth1,c("methadon","methodon"),0.01)

freqterms <- findFreqTerms(dtmmeth1, lowfreq = 400)
print(freqterms)
```
* create document term matrix
* find associations
* find freqent terms


```{r meth1new, echo=FALSE}
myStopwords <- c("methadon","methodon","methados","dolophin")
meth_corpus1_new <- tm_map(meth_corpus1, removeWords, myStopwords)


wordcloud(meth_corpus1_new, min.freq=400, max.words=15, scale=c(3,.1), random.order=FALSE, colors=brewer.pal(12, "Set3"))
```
* add new stop words
* generate word cloud


# Fentanyl Tweet Set 
```{r fenttext1, echo=FALSE}
fent_text1 <- tweet_text1[c(grep(fentanyl, tweet_text1,ignore.case=TRUE))]
fent_text2 <- tweet_text2[c(grep(fentanyl, tweet_text2,ignore.case=TRUE))]

fent_text <- append(fent_text1,fent_text2)
```
* load fentanyl data

```{r prefent1, echo=FALSE}
#preprocess subtext
fent_corpus1 <- VCorpus(VectorSource(fent_text))

#preprocessing: lowercase
fent_corpus1 <- tm_map(fent_corpus1, content_transformer(tolower))
#remove stop words 
fent_corpus1 <- tm_map(fent_corpus1, removeWords, stopwords("en"))
fent_corpus1 <- tm_map(fent_corpus1, removeWords, stopwords("ca"))
fent_corpus1 <- tm_map(fent_corpus1, removeWords, stopwords("es"))
fent_corpus1 <- tm_map(fent_corpus1, removeWords, stopwords("fr"))

#preprocessing: remove punctuation
fent_corpus1 <- tm_map(fent_corpus1, removePunctuation)
#preprocessing:remove numbers
fent_corpus1 <- tm_map(fent_corpus1, removeNumbers)
#preprocessing: remove stems
fent_corpus1 <- tm_map(fent_corpus1, stemDocument)
```
* preprocessing


```{r dtmfent1, echo=FALSE}
dtmfent1 <-removeSparseTerms(DocumentTermMatrix(fent_corpus1),0.995)
inspect(dtmfent1)
findAssocs(dtmfent1,c("fentanyl","fentanil","fentan"),0.01)

freqterms <- findFreqTerms(dtmfent1, lowfreq = 5000)
print(freqterms)
```
* create doc term matrix
* find associations
* find frequent terms


```{r fentnew1, echo=FALSE}
myStopwords <- c("fentanyl","fentanil")
fent_corpus1_new <- tm_map(fent_corpus1, removeWords, myStopwords)


wordcloud(fent_corpus1_new, min.freq=3000, max.words=15, scale=c(3,.1), random.order=FALSE, colors=brewer.pal(12, "Set3"))
```
* add new stop words
* generate word cloud


# Fentanyl Bigrams
```{r token, echo=FALSE}
BigramTokenizer <-
  function(x)
    unlist(lapply(ngrams(words(x), 2), paste, collapse = "_"), use.names = FALSE)
```
* Bigram Tokenizer
```{r fentbigram1, echo=FALSE}
dtm_bigramfent1 <- removeSparseTerms(TermDocumentMatrix(fent_corpus1, control = list(tokenize = BigramTokenizer)),0.995)
```
* create bigram tdm

```{r fentbigramplot1, echo=FALSE}
y <- as.matrix(dtm_bigramfent1 )
ysum <- rowSums(y)

par(mar=c(10,6,4,1)+.1)
barplot(sort(ysum,decreasing=TRUE)[1:10],las=2,xlab ="Bigrams", 
        main = "Frequency of Bigrams in Fentanyl Tweets", mgp=c(8,.3,0))

title(ylab="Frequency", line=3, cex.lab=1.2)
```
* plot frequency

# Suboxone Bigrams 
```{r subbi1, echo=FALSE}
dtm_bigramsub1 <- removeSparseTerms(TermDocumentMatrix(subcorpus_corpus1, control = list(tokenize = BigramTokenizer)),0.995)
```
* create bigram tdm

```{r subbiplot1, echo=FALSE}
y <- as.matrix(dtm_bigramsub1 )
ysum <- rowSums(y)

par(mar=c(10,6,4,1)+.1)
barplot(sort(ysum,decreasing=TRUE)[1:10],las=2,xlab ="Bigrams", 
        main = "Frequency of Bigrams in Suboxone Tweets", mgp=c(8,.3,0))

title(ylab="Frequency", line=3, cex.lab=1.2)
```
* plot frequency

# Methadone Bigrams Tweet Set 1
```{r methbi1, echo=FALSE}
dtm_bigrammeth1 <- removeSparseTerms(TermDocumentMatrix(meth_corpus1, control = list(tokenize = BigramTokenizer)),0.995)
```
* create bigram tdm

```{r methbiplot1, echo=FALSE}
y <- as.matrix(dtm_bigrammeth1 )
ysum <- rowSums(y)

par(mar=c(10,6,4,1)+.1)
barplot(sort(ysum,decreasing=TRUE)[1:10],las=2,xlab ="Bigrams", 
        main = "Frequency of Bigrams in Methadone Tweets", mgp=c(8,.3,0))

title(ylab="Frequency", line=3, cex.lab=1.2)
```
* plot frequency


# Methadone Time Series
```{r methadonetime, echo=FALSE}
meth_text1 <- tweets_set1[c(grep(methadone, tweets_set1$text,ignore.case=TRUE)),]
meth_text2 <- tweets_set2[c(grep(methadone, tweets_set2$text,ignore.case=TRUE)),]
meth_text <- rbind(meth_text1,meth_text2)
meth_text$date <- as.Date(meth_text$date, format = "%Y-%m-%d")
summary(meth_text)
```
* load and combine tweet sets
* format date

```{r mtimeplot, echo=FALSE}
#time frequency
freqs <- aggregate(meth_text$date, by=list(meth_text$date), FUN=length) 
freqs$date <- as.Date(freqs$Group.1, format="%Y-%m-%d") 
ggplot(freqs, aes(x=date, y=x)) + geom_line() + ylab("Frequency") + xlab("Date (Nov 2019 - Apr 2020) ") + theme_minimal() + labs( title ="Frequency of Methdone-related Tweets by Date")
```
* plot frequency

# Suboxone Time Series
```{r suboxonetime, echo=FALSE}
sub_text1 <- tweets_set1[c(grep(sub, tweets_set1$text,ignore.case=TRUE)),]
sub_text2 <- tweets_set2[c(grep(sub, tweets_set2$text,ignore.case=TRUE)),]
sub_text <- rbind(sub_text1,sub_text2)
sub_text$date <- as.Date(sub_text$date, format = "%Y-%m-%d")
summary(sub_text)
```
* load and combine tweet sets
* format date

```{r stimeplot, echo=FALSE}
#time frequency
freqs <- aggregate(sub_text$date, by=list(sub_text$date), FUN=length) 
freqs$date <- as.Date(freqs$Group.1, format="%Y-%m-%d") 
ggplot(freqs, aes(x=date, y=x)) + geom_line() + ylab("Frequency") + xlab("Date (Nov 2019 - Apr 2020) ") + theme_minimal() + labs( title ="Frequency of Suboxone-related Tweets by Date")
```
* plot frequency

# Fentanyl Time Series
```{r fentanyltime, echo=FALSE}
fenttext1 <- tweets_set1[c(grep(fentanyl, tweets_set1$text,ignore.case=TRUE)),]
fenttext2 <- tweets_set2[c(grep(fentanyl, tweets_set2$text,ignore.case=TRUE)),]
fenttext<- rbind(fenttext1,fenttext2)
fenttext$date <- as.Date(fenttext$date, format = "%Y-%m-%d")
```
* load and combine tweet sets
* format date

```{r ftimeplot, echo=FALSE}
#time frequency
freqs <- aggregate(fenttext$date, by=list(fenttext$date), FUN=length) 
freqs$date <- as.Date(freqs$Group.1, format="%Y-%m-%d") 
 ggplot(freqs, aes(x=date, y=x)) + geom_line() + ylab("Frequency") + xlab("Date (Nov 2019 - Apr 2020) ") + theme_minimal() + labs( title ="Frequency of Fentanyl-related Tweets by Date")
```
* plot frequency

