---
title: "Assignment 2"
output: html_document
date: "2023-11-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(caTools)
library(rfUtilities)
library(tm)
library(mlapi)
library(e1071)
library(SnowballC)
library(caret)
library(ggplot2)
```

* load dataset
```{r}
annotated_dataset <- read.csv('Datasets/TADA_Annotated_data.csv')
unlabeled_dataset <- read.csv('Datasets/TADA_unlabeled_data.csv')
```

* load and preprocess text
```{r}
all_texts <-annotated_dataset$text
all_texts_corpus <- VCorpus(VectorSource(all_texts))
all_texts_corpus <- tm_map(all_texts_corpus, content_transformer(function(x) iconv(enc2utf8(x), sub = "byte")))
all_texts_corpus <- tm_map(all_texts_corpus, content_transformer(tolower))
all_texts_corpus <- tm_map(all_texts_corpus, removePunctuation)
all_texts_corpus <- tm_map(all_texts_corpus, removeWords,stopwords("english"))
all_texts_corpus <- tm_map(all_texts_corpus, stemDocument)
length(all_texts_corpus)
```

```{r}
NLP_tokenizer <- function(x) {
  unlist(lapply(ngrams(words(x), 1:3), paste, collapse = "_"), use.names = FALSE)
}
n_gram_corpus <- tm_map(all_texts_corpus,content_transformer(NLP_tokenizer))
```

* view corpus and content
```{r}
length(n_gram_corpus)
n_gram_corpus[[2]]$content
```

* split the sets 
```{r}
set.seed(1234)
split <- sample.split(annotated_dataset$class,SplitRatio = 0.8)
training_corpus <- subset(n_gram_corpus, split==TRUE)
training_classes <- subset(annotated_dataset$class, split==TRUE)

eval_corpus <- subset(n_gram_corpus, split==FALSE)
eval_classes <- subset(annotated_dataset$class, split==FALSE)

split2 <- sample.split(eval_classes,SplitRatio = 0.5)
validation_corpus <- subset(eval_corpus, split2==TRUE)
validation_classes <- subset(eval_classes, split2==TRUE)
  
test_corpus <- subset(eval_corpus, split2==FALSE)
test_classes <- subset(eval_classes, split2==FALSE)
```

* create document-term matrix for the training set and remove sparse n-grams.
```{r}
training_dtm <- DocumentTermMatrix(training_corpus)
training_dtm_sparse <- removeSparseTerms(training_dtm,0.995)
```

* generate document-term matrix for validation set and test set using columns from training set 
```{r}
validation_dtm_sparse <- DocumentTermMatrix(validation_corpus, list(dictionary=colnames(training_dtm_sparse)))
test_dtm_sparse <- DocumentTermMatrix(test_corpus, list(dictionary=colnames(training_dtm_sparse)))
```

* convert to data frames 
```{r}
training_df <- as.data.frame(as.matrix(training_dtm_sparse))
validation_df <- as.data.frame(as.matrix(validation_dtm_sparse))
test_df <- as.data.frame(as.matrix(test_dtm_sparse))
colnames(training_df) <- make.names(colnames(training_df))
colnames(validation_df) <- make.names(colnames(validation_df))
colnames(test_df) <- make.names(colnames(test_df))
training_df$class <- training_classes
training_df$class <- as.factor(training_df$class)
```


* SVM- radial kernel
```{r}
i = 1
while  (i <= 64){ 
  trained_model <- svm(class ~., data=training_df,cost=i,kernel = "radial")
  predictions <- predict(trained_model, newdata=validation_df)
  print(i)
  print(accuracy(validation_classes,predictions))
i = i*2
}
```

* SVM cost of 4
```{r}
best_trained_model_svm <- svm(class ~., data=training_df,cost=4,kernel = "radial")
best_predictions_svm <- predict(best_trained_model_svm, newdata=test_df)
```

*SVM accuracy and confusion matrix
```{r}
print(accuracy(test_classes,best_predictions_svm))
confusionMatrix(best_predictions_svm,as.factor(test_classes),mode="everything")
```

* k-nearest neighbor
```{r}
i = 1
while  (i <= 64){ 
  trained_model <- gknn(class ~., data=training_df,k = i,method = "Manhattan")
  predictions <- predict(trained_model, newdata=validation_df)
  print(i)
  print(accuracy(validation_classes,predictions))
i = i*2
}
```

* knn cost of 8
```{r}
best_trained_model_knn <- gknn(class ~., data=training_df,k = 8, method = "Manhattan")
best_predictions_knn <- predict(best_trained_model_knn, newdata=test_df)
```

* knn accuracy and confusion matrix
```{r}
print(accuracy(test_classes,best_predictions_knn))
confusionMatrix(best_predictions_knn,as.factor(test_classes),mode="everything")
```

* naive Bayes
```{r}
trained_model <- naiveBayes(class ~., data=training_df)
predictions_nb <- predict(trained_model, newdata=validation_df,type = "class")
```

* naive Bayes accuracy and confusion matrix
```{r}
print(accuracy(validation_classes,predictions))
confusionMatrix(predictions_nb,as.factor(test_classes),mode="everything")
```



*preprocess unlabeled data
```{r}
all_texts_unlabeled <-unlabeled_dataset$text
all_texts_corpus_unlabeled <- VCorpus(VectorSource(all_texts_unlabeled))
all_texts_corpus_unlabeled <- tm_map(all_texts_corpus_unlabeled, content_transformer(function(x) iconv(enc2utf8(x), sub = "byte")))
all_texts_corpus_unlabeled <- tm_map(all_texts_corpus_unlabeled, content_transformer(tolower))
all_texts_corpus_unlabeled <- tm_map(all_texts_corpus_unlabeled, removePunctuation)
all_texts_corpus_unlabeled <- tm_map(all_texts_corpus_unlabeled, removeWords,stopwords("english"))
all_texts_corpus_unlabeled <- tm_map(all_texts_corpus_unlabeled, stemDocument)
length(all_texts_corpus_unlabeled)
```

* create bigrams
```{r}
n_gram_corpus_unlabeled <- tm_map(all_texts_corpus_unlabeled,content_transformer(NLP_tokenizer))
```

* generate DTM
```{r}
unlabeled_dtm <- DocumentTermMatrix(n_gram_corpus_unlabeled,list(dictionary=colnames(training_dtm_sparse)))
```

* convert to data frames 
```{r}
unlabeled_df <- as.data.frame(as.matrix(unlabeled_dtm))
colnames(unlabeled_df) <- make.names(colnames(unlabeled_df))
```

* classify unlabeled data
```{r}
predictions_unlabeled_svm <- predict(best_trained_model_svm, newdata=unlabeled_df)
```

```{r}
summary(predictions_unlabeled_svm)
```

* map labels to data
```{r}
new_labeled_data <- cbind(unlabeled_dataset,predictions_unlabeled_svm)
colnames(new_labeled_data)[colnames(new_labeled_data) == "predictions_unlabeled_svm"] = "class"
```

```{r}
cityA_set <- new_labeled_data[new_labeled_data$city == "A",]
cityB_set <- new_labeled_data[new_labeled_data$city == "B",]
```

```{r}
pop_A <- 500000
pop_B <- 10000
```

```{r}
table(cityA_set$class)
table(cityB_set$class)
```


```{r}
F_set <- new_labeled_data[new_labeled_data$gender_id == "F",]
F_set<- F_set %>%
  mutate(class = recode(class, "0" = "Non-medical use", "1" = "Consumption", "2" = "Information/mention", "3" = "Unrelated"))
M_set <- new_labeled_data[new_labeled_data$gender_id == "M",]
M_set<- M_set %>%
  mutate(class = recode(class, "0" = "Non-medical use", "1" = "Consumption", "2" = "Information/mention", "3" = "Unrelated"))
summary(F_set)
summary(M_set)
```

```{r}
table(F_set$class)
table(M_set$class)
```

```{r}
ggplot(F_set, aes(x = class)) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(title = "Frequency Distribution", x = "Categories", y = "Frequency")
```

```{r}
ggplot(M_set, aes(x = class)) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(title = "Frequency Distribution", x = "Categories", y = "Frequency")
```

* city A tweet distribution
```{r}
frequency_tableA <- table(cityA_set$class)
adjusted_frequencyA<- frequency_tableA / pop_A

# Create a data frame for plotting
plot_data <- data.frame(
  Class = names(adjusted_frequencyA),
  Frequency = as.numeric(adjusted_frequencyA)
)

plot_data<- plot_data %>%
  mutate(Class = recode(Class, "0" = "Non-medical use", "1" = "Consumption", "2" = "Information/mention", "3" = "Unrelated"))

# Create a bar plot
ggplot(plot_data, aes(x = Class, y = Frequency)) +
  geom_bar(stat = "identity", fill = "white", color = "black") +
   geom_text(aes(label = sprintf("%.5f", Frequency)), vjust = -0.5) +
  labs(title = "Population-Adjusted Distribution of Tweets for City A",
       x = "Class",
       y = "Population-Adjusted Frequency")
```


*city B tweet distribution
```{r}
frequency_tableB <- table(cityB_set$class)
adjusted_frequencyB<- frequency_tableB / pop_B

# Create a data frame for plotting
plot_data <- data.frame(
  Class = names(adjusted_frequencyB),
  Frequency = as.numeric(adjusted_frequencyB)
)

plot_data<- plot_data %>%
  mutate(Class = recode(Class, "0" = "Non-medical use", "1" = "Consumption", "2" = "Information/mention", "3" = "Unrelated"))

# Create a bar plot
ggplot(plot_data, aes(x = Class, y = Frequency)) +
  geom_bar(stat = "identity", fill = "white", color = "black") +
   geom_text(aes(label = sprintf("%.5f", Frequency)), vjust = -0.5) +
  labs(title = "Population-Adjusted Distribution¸ of Tweets for City B",
       x = "Class",
       y = "Population-Adjusted Frequency")
```


